Optimal Algorithms for Dynamic Balanced Graph Partitioning

### 2020-05-14 paper/disc20clustering.tex
one line LB formula
minor

	  Next, we argue that a partition exists when the process terminates.
	 This implies that a feasible partition exists after any earlier step as well.
%	 Assume for contradiction that the partition is not feasible.	 
	 Since there are at least $\ell+1$ isolated nodes left,
	 there must be two isolated nodes $x^*$ and $y^*$,
	  with the same cluster of origin,
	 i.e., $I(\{x^*\}) = I(\{y^*\})$.
	 Consider the partition $P^*$
	  obtained from the initial partition after swapping $x_0$ and $y_0$
	 with $x^*$ and $y^*$ (respectively).
	 In this partition,
	 the ground set $\{x_0,y_0\}$ is collocated in the cluster $I(\{x^*,y^*\})$.
	 Note that after the first request $\{x_0,y_0\}$,
	   we issue requests only between nodes that have the same cluster of origin and
	   all these nodes are collocated in $P^*$.
	 Therefore all ground sets constructed so far are collocated in  $P^*$,
	 i.e., $P^*$ is a feasible partition.
%	 Hence, there are at most $\ell$ isolated nodes after the termination.

@@

	\ALG performs at least one swap each time a~ground set grows,
	i.e., at each step $i$.
%	In other words,
%	some ground set grows by one node at each of these steps.
	Consider any ground set $C^* \neq B$ after the termination.
	This ground set has grown exactly $|C^*|-1$ times until the the termination. 
	Let $\mathcal{S}$ be the set of all ground sets after the process terminates.
	Thus, $\mathcal{S}$ includes ground sets $B$,
	$\{x_0,y_0\}$,
	and (up to) $\ell+2$ singleton ground sets.
	Among the remaining ground sets in $\mathcal{S}$,
	no two ground sets have the same origin.
	Otherwise,
   the smaller ground set is either a singleton,
   which contradicts the bound $\ell+2$ on the number of singletons,
	or we have joined nodes to it at some step,
	contradicting our choice of the largest $C_i$ at step $i$.
	Hence, every  other ground set in
	$\mathcal{S}$ has grown \emph{maximally}.
	There are at most $\ell-1$ maximally grown ground sets,
	one per cluster,
	 excluding the cluster containing $B$. 
	Therefore,  $|\mathcal{S}| \leq 1 + 1 + (\ell+2)  + (\ell-1) = 2\ell+3$.
	Note that among all non-singleton ground sets in $\mathcal{S}$,
	 only $B$ does not grow during the process.
	 Thus,
    the total number of times that 
	a ground set in $\mathcal{S}$  has grown is
	\begin{align*}
	\sum_{C^* \in \mathcal{S} } (|C^*|-1) - (k-1) &= 
	\sum_{C^* \in \mathcal{S} } |C^*| - 
		\sum_{C^* \in \mathcal{S} } 1 - (k-1) \\
	&\geq  k  \ell - (2\ell+3) - (k-1) =  (k-2)(\ell-1) - 4,
	\end{align*}
	which bounds the number of swaps performed by \ALG.
	The competitive ratio is then
	$\ALG / \OPT \geq ((k-2)(\ell-1) - 4) / 2$.
	%%
	\begin{comment}
	\ALG performs at least one swap  every time   a~ground set is constructed.
	Each expansion reduces the number of isolated nodes by $1$ or $2$.
	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still singleton; this occurs at most $\ell$ times.
	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
	\end{comment}
	%%
\end{proof}

>>>>

	  Next, we argue that a feasible partition exists when the process terminates.
	 This implies that a feasible partition exists after any earlier step as well.
%	 Assume for contradiction that the partition is not feasible.	 
	 Since there are at least $\ell+1$ isolated nodes left,
	 there must be two isolated nodes $x^*$ and $y^*$,
	  with the same cluster of origin,
	 i.e., $I(\{x^*\}) = I(\{y^*\})$.
	 Consider the partition $P^*$
	  obtained from the initial partition after swapping $x_0$ and $y_0$
	 with $x^*$ and $y^*$ (respectively).
	 In this partition,
	 the ground set $\{x_0,y_0\}$ is collocated in the cluster $I(\{x^*,y^*\})$.
	 Note that after the first request $\{x_0,y_0\}$,
	   we issue requests only between nodes that have the same cluster of origin and
	   all these nodes are collocated in $P^*$.
	 Therefore all ground sets constructed so far are collocated in  $P^*$,
	 i.e., $P^*$ is a feasible partition.
%	 Hence, there are at most $\ell$ isolated nodes after the termination.

@@

	\ALG performs at least one swap each time a~ground set grows,
	i.e., at each step $i$.
%	In other words,
%	some ground set grows by one node at each of these steps.
	Consider any ground set $C^* \neq B$ after the termination.
	This ground set has grown exactly $|C^*|-1$ times until the the termination. 
	Let $\mathcal{S}$ be the set of all ground sets after the process terminates.
	Thus, $\mathcal{S}$ includes ground sets $B$,
	$\{x_0,y_0\}$,
	and (up to) $\ell+2$ singleton ground sets.
	Among the remaining ground sets in $\mathcal{S}$,
	no two ground sets have the same origin.
	Otherwise,
   the smaller ground set is either a singleton,
   which contradicts the bound $\ell+2$ on the number of singletons,
	or we have joined nodes to it at some step,
	contradicting our choice of the largest $C_i$ at step $i$.
	Hence, every  other ground set in
	$\mathcal{S}$ has grown \emph{maximally}.
	There are at most $\ell-1$ maximally grown ground sets,
	one per cluster,
	 excluding the cluster containing $B$. 
	Therefore,  $|\mathcal{S}| \leq 1 + 1 + (\ell+2)  + (\ell-1) = 2\ell+3$.
	Note that among all non-singleton ground sets in $\mathcal{S}$,
	 only $B$ does not grow during the process.
	 Thus,
    the total number of times that 
	a ground set in $\mathcal{S}$  has grown is
	$	\sum_{C^* \in \mathcal{S} } (|C^*|-1) - (k-1) $
	\begin{align*}
	=\sum_{C^* \in \mathcal{S} } |C^*| - 
		\sum_{C^* \in \mathcal{S} } 1 - (k-1) 
	\geq  k  \ell - (2\ell+3) - (k-1) =  (k-2)(\ell-1) - 4,
	\end{align*}
	which bounds the number of swaps performed by \ALG.
	The competitive ratio is then
	$\ALG / \OPT \geq ((k-2)(\ell-1) - 4) / 2$.
	%%
	\begin{comment}
	\ALG performs at least one swap  every time   a~ground set is constructed.
	Each expansion reduces the number of isolated nodes by $1$ or $2$.
	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still singleton; this occurs at most $\ell$ times.
	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
	\end{comment}
	%%
\end{proof}
